Preprocessing and exploration of the narwhal data for the analysis with smoothSDE

### 2025-03-17 R/preprocessing/narwhals/data_preprocessing.Rmd
update preprocessing of borders and exploratory plots

## Import necessary packages

@@

## Choose directory where the plots should be saved

@@

#add time since tagging column

TimeFunction <- function(n, dt = 1){
  ## n: Number of observations. Positive integer
  ## dt: Time step between observations in seconds
  Time <- (1:n)*(dt/3600) ## Time in hours since tagging
}

@@

#animated map with gganimate
animate_map=milne_land_map + 
  geom_line(dataprep[,c("ID","Long","Lat","time")], 
            mapping=aes(Long, Lat, col = ID),size= 1,alpha=0.5) +
  scale_color_viridis_d(labels=c("A1","A2","A3","A4","A5","A6"))+transition_reveal(time)

animation=animate(animate_map,renderer = gifski_renderer())
anim_save(path=plots_folder,filename="all_narwhals_tracks.gif", 
          animation = animation,width=10,height=5)
```

@@

scoresby_sound_utm=st_transform(scoresby_sound_utm,crs= "+init=EPSG:32626 +units=km")
```

@@

```{r plot coastline and land}
plot_land=ggplot()+geom_sf(data=scoresby_sound_utm$geometry,fill="grey",col="brown")
```

```{r display land , fig.cap="Land polygons"}
ggplotly(plot_land)
```

@@

## Points on land

@@

```{=tex}
A lot of points are indeed still very close to land. We think that this means the shoreline is still not accurate. To solve the problem, we translate the polygon vertices 200m inward. We compute new distances to shore based on the translated polygon.

@@

@@

#get updated land polygons
shrunk_scoresby_sound_utm<-st_read(file.path(preprocessed_greenland_data_path,"shrunk_updated_scoresby_sound_utm.shp"))

shrunk_scoresby_sound_utm <- st_transform(shrunk_scoresby_sound_utm, crs = "+init=EPSG:32626 +units=km")

@@

```{r compute updated translated distance to shore}
res_shrunk_land=nearest_points(positions=narwhal_positions,border=shrunk_scoresby_sound_utm)

cat("Percentage of points on land with distance from shrunk updated osm land polygons : \n",
    length(which(res_shrunk_land$min_distances==0))/length(dataprep$time)*100,"\n")

@@

df=data.frame(DistanceShore=res_shrunk_land$min_distances[res_shrunk_land$min_distances<1 & res_shrunk_land$min_distances>0  ])
histo=ggplot(df, aes(x =DistanceShore)) +
  geom_histogram(fill="cyan",bins=100,color="black")+xlab("Computed distance to shore (km)")

@@

#keep points calculated from land polygons
nearest_points=res_updated_land$nearest_points
min_distances=res_updated_land$min_distances

#add columns px and py with coordinates of the nearest point on the shore line
dataprep$px=nearest_points[,1]
dataprep$py=nearest_points[,2]

# add columns with difference between narwhal position and nearest point on the shoreline
dataprep$nx=dataprep$x-dataprep$px
dataprep$ny=dataprep$y-dataprep$py

#replace distance to shore with distances calculated from land polygons (in km)
dataprep$DistanceShore=min_distances

#same for shrunk polygons
nearest_points_shrunk=res_shrunk_land$nearest_points
min_distances_shrunk=res_shrunk_land$min_distances

#add columns px and py with coordinates of the nearest point on the shore line
dataprep$px_translated=nearest_points_shrunk[,1]
dataprep$py_translated=nearest_points_shrunk[,2]

# add columns with difference between narwhal position and nearest point on the shoreline
dataprep$nx_translated=dataprep$x-dataprep$px_translated
dataprep$ny_translated=dataprep$y-dataprep$py_translated
dataprep$DistanceShore_translated=min_distances_shrunk
```

@@

```{r}
#add Ishore covariate based on 1/distance in km
dataprep$Ishore <- ExposureFunction(X = dataprep$DistanceShore)

#add Ishore covariate based on 1/distance in km
dataprep$Ishore_translated <- ExposureFunction(X = dataprep$DistanceShore_translated)

@@

dataprep$AngleShore=rep(NA,n)
dataprep$AngleVelocities=rep(NA,n)
dataprep$theta=rep(NA,n)
dataprep$theta_translated=rep(NA,n)
dataprep$VelocityPhase=rep(NA,n)

@@

  #matrix of normal vectors
  normal=as.matrix(sub_data[2:n_sub,c("nx","ny")])
  normal_translated=as.matrix(sub_data[2:n_sub,c("nx_translated","ny_translated")])

  #angle between velocity and normal vector
  theta=signed_angle(vexp_df,normal)
  theta_translated=signed_angle(vexp_df,normal_translated)

@@

  #deviation from trajectory parallel to coastline
  phi=ifelse(theta > 0, theta - pi/2, theta + pi/2)

  #adjust lengths by adding NA when necessary
  angle_vel=c(NA,angle_vel,NA)
  phi=c(NA,phi)
  theta=c(NA,theta)
  theta_translated=c(NA,theta_translated)
  vel_phase=c(NA,vel_phase,NA)

  dataprep[sub_ind,"AngleShore"]=phi
  dataprep[sub_ind,"AngleVelocities"]=angle_vel
  dataprep[sub_ind,"theta"]=-theta
  dataprep[sub_ind,"theta_translated"]=-theta_translated
  dataprep[sub_ind,"VelocityPhase"]=vel_phase
}

@@

plot_AngleShore=ggplot()+geom_line(data=dataprep[dataprep$DistanceShore>0.05,],aes(time,AngleShore,col=ID))+
  geom_rect(data = regions, aes(xmin = start_time, xmax = end_time, ymin = -Inf, ymax = Inf),alpha = 0.3,col="lightblue")+
  facet_grid(ID ~.,labeller=as_labeller(c("Asgeir"="A1","Frederik"="A2","Helge18"="A3","Kyrri"="A4","Nemo"="A5","Siggi"="A6")))+
  scale_color_viridis_d(guide="none")+theme(legend.position="none")+xlab("Time (h)")+ylab("Empirical deviation from coastline")+
  theme_minimal()

@@

plot_AngleNormal=ggplot()+geom_line(data=dataprep[dataprep$DistanceShore>0.05,],aes(time,theta,col=ID))+
  geom_rect(data = regions, aes(xmin = start_time, xmax = end_time, ymin = -Inf, ymax = Inf),alpha = 0.3,col="lightblue")+
  facet_grid(ID ~.,labeller=as_labeller(c("Asgeir"="A1","Frederik"="A2","Helge18"="A3","Kyrri"="A4","Nemo"="A5","Siggi"="A6")))+
  scale_color_viridis_d(guide="none")+theme(legend.position="none")+xlab("Time (h)")+ylab("Angle between empirical velocity and normal vector")+ theme_minimal()

ggsave(path=plots_folder,"plot_horizontal_speed.png", plot = plot_speed)
ggsave(path=plots_folder,"plot_AngleShore.png", plot = plot_AngleShore)
ggsave(path=plots_folder,"plot_AngleVelocities.png", plot = plot_AngleVelocities)
ggsave(path=plots_folder,"plot_AngleNormal.png", plot = plot_AngleNormal)

@@

```{r display plot_AngleShore , fig.cap="",warning=FALSE}
plot_AngleShore
plot_AngleVelocities
```

@@

AngleShore_histo=ggplot(dataprep[dataprep$DistanceShore>0.05,], aes(x = AngleShore,col=ID)) +
  geom_histogram(fill="white",bins=50)+facet_grid(ID ~.,labeller=as_labeller(c("Asgeir"="A1","Frederik"="A2","Helge18"="A3","Kyrri"="A4","Nemo"="A5","Siggi"="A6")))+
  scale_color_viridis_d(guide="none")+
  theme(legend.position="top")+xlab("Empirical deviation from coastline (rad)")+ theme_minimal()

@@

AngleNormal_histo=ggplot(dataprep[dataprep$DistanceShore>0.05,], aes(x = theta,col=ID)) +
  geom_histogram(fill="white",bins=50)+facet_grid(ID ~.,labeller=as_labeller(c("Asgeir"="A1","Frederik"="A2","Helge18"="A3","Kyrri"="A4","Nemo"="A5","Siggi"="A6")))+
  scale_color_viridis_d(guide="none")+
  theme(legend.position="top")+xlab("Angle between empirical velocity and normal vector (rad) ")+ theme_minimal()

@@

ggsave(path=plots_folder,"AngleShore_histo.png", plot = AngleShore_histo)
ggsave(path=plots_folder,"AngleVelocities.png", plot = AngleVelocities_histo)
ggsave(path=plots_folder,"AngleNormal.png", plot = AngleNormal_histo)
ggsave(path=plots_folder,"VelocityPhase.png", plot = VelocityPhase_histo)
```

```{r display AngleShore_histo, fig.cap="",warning=FALSE}
AngleShore_histo
```

@@

plot_AngleShoreDistance=ggplot()+geom_point(data=dataprep,aes(DistanceShore,AngleShore,col=ID))+
  facet_grid(ID ~.,labeller=as_labeller(c("Asgeir"="A1","Frederik"="A2","Helge18"="A3","Kyrri"="A4","Nemo"="A5","Siggi"="A6")))+
  scale_color_viridis_d(guide="none")+theme(legend.position="none")+xlab("Distance to shore (km)")+ylab("Empirical deviation from coastline")

ggsave(path=plots_folder,"plot_AngleShore_DistanceShore.png", plot = plot_AngleShoreDistance)

@@

# Plot smoothed AngleSHore
plot_AngleShoreDistanceSmoothed <-  ggplot()+ geom_smooth(data=dataprep,aes(DistanceShore,abs(AngleShore),col=ID))+
  facet_grid(ID ~.,labeller=as_labeller(c("Asgeir"="A1","Frederik"="A2","Helge18"="A3","Kyrri"="A4","Nemo"="A5","Siggi"="A6")))+
  scale_color_viridis_d(guide="none")+theme(legend.position="none")+xlab("Distance to shore (km)")+ylab("Empirical deviation from coastline")

@@

```{r display plot_AngleShoreDistanceSmoothed, fig.cap=""}
plot_AngleShoreDistanceSmoothed
```

@@

```{r plot empirical velocities and deviation angles without points on land}
# Plot smoothed AngleSHore
plot_AngleShoreDistanceSmoothed <-  ggplot()+
  geom_smooth(data=dataprep[dataprep$DistanceShore>0,],aes(DistanceShore,abs(AngleShore),col=ID))+
  facet_grid(ID ~.,labeller=as_labeller(c("Asgeir"="A1","Frederik"="A2","Helge18"="A3","Kyrri"="A4","Nemo"="A5","Siggi"="A6")))+
  scale_color_viridis_d(guide="none")+
  theme(legend.position="none")+xlab("Distance to shore (km)")+
  ylab("Empirical deviation from coastline")

@@

```{r distance shore histogram}
Dshore_histo=ggplot(dataprep, aes(x = DistanceShore)) +
  geom_histogram(fill="white",color="black",bins=50)+
  facet_grid(ID ~.,labeller=as_labeller(c("Asgeir"="A1","Frederik"="A2","Helge18"="A3","Kyrri"="A4","Nemo"="A5","Siggi"="A6")))+
  theme(legend.position="none")+xlab("Distance to shore (km)")

ggsave(path=plots_folder,"Dshore_histo.png", plot = Dshore_histo)
```

```{r display Dshore_histo, fig.cap="Histograms of distance to shore"}
Dshore_histo
```
## Plot distance to ship histogram for each narwhal

@@

Dship_histo=ggplot(dataprep, aes(x = Dist_Ship)) +
  geom_histogram(fill="white",col="black",bins=30)+
  facet_grid(ID ~.,labeller=as_labeller(c("Asgeir"="A1","Frederik"="A2","Helge18"="A3","Kyrri"="A4","Nemo"="A5","Siggi"="A6")))+
  theme(legend.position="none")+xlab("Distance to ship (km)")

@@

```{r}
D_low=0.07
D_up=3
df=cropped_data
df[df$DistanceShore<D_low,"Ishore"]=1/D_low
df[df$DistanceShore>D_up,"Ishore"]=0

pIshore=ggplot(df,aes(time,Ishore,col=ID))+ geom_line(color="black")+facet_grid(ID ~.,labeller=as_labeller(c("Asgeir"="A1","Frederik"="A2","Helge18"="A3","Kyrri"="A4","Nemo"="A5","Siggi"="A6")))+
  theme(legend.position="none")+xlab("Time (h)")+ylab("Inverse distance to shore")

ggsave(path=plots_folder,"realIshore_through_time.png", plot=pIshore)
```

@@

>>>>

## Imports

@@

## Create plots directory

@@

#add time since tagging column
TimeFunction <- function(n, dt = 1){
  ## n: Number of observations. Positive integer
  ## dt: Time step between observations in seconds
  Time <- (1:n)*(dt/3600) ## Time in hours since tagging
}

@@

#animated map with gganimate
animate_map <- milne_land_map + 
  geom_point(data = dataprep[,c("ID","Long","Lat","time")], 
             mapping = aes(Long, Lat, shape = ID, col = ID), 
             size = 4, alpha = 1) +
  geom_path(data = dataprep[,c("ID","Long","Lat","time")], 
            mapping = aes(Long, Lat, col = ID), 
            size = 1, alpha = 0.5) +
  scale_color_viridis_d(name = "ID",  # Give a common legend title
                        labels = c("A1","A2","A3","A4","A5","A6")) + 
  scale_shape_manual(name = "ID",  # Same name as color to merge legends
                     values = c(0, 1, 2, 3, 4, 5),  
                     labels = c("A1","A2","A3","A4","A5","A6")) +  
  guides(color = "legend", shape = "legend") +  # Ensure single legend
  transition_reveal(time) +
  labs(subtitle = "Time: {round(frame_along,1)} hours")

animate_map

animation=animate(animate_map,nframes = 100, fps = 5,renderer = gifski_renderer())
anim_save(path=plots_folder,filename="all_narwhals_tracks.gif", 
          animation = animation,width=10,height=5)
```

@@

scoresby_sound_utm<-st_transform(scoresby_sound_utm,crs= "+init=EPSG:32626 +units=km")
```

@@

```{r plot coastline and land}
plot_land=ggplot()+geom_sf(data=scoresby_sound_utm$geometry,fill="grey",col="brown")
ggplotly(plot_land)
```

@@

@@

```{=tex}
A lot of points are indeed still very close to land. We think that this means the shoreline is still not accurate.
To solve the problem, we translate the polygon vertices 100m, 200m and 300m inward. 
We compute new distances to shore based on the translated polygon.

@@

## Compute distances from shrunk land polygons

@@

#get shrunk land polygons
shrunk100_scoresby_sound_utm<-st_read(file.path(preprocessed_greenland_data_path,"shrunk100_scoresby_sound_utm.shp"))
shrunk100_scoresby_sound_utm <- st_transform(shrunk100_scoresby_sound_utm, crs = "+init=EPSG:32626 +units=km")

shrunk200_scoresby_sound_utm<-st_read(file.path(preprocessed_greenland_data_path,"shrunk200_scoresby_sound_utm.shp"))
shrunk200_scoresby_sound_utm <- st_transform(shrunk200_scoresby_sound_utm, crs = "+init=EPSG:32626 +units=km")

shrunk300_scoresby_sound_utm<-st_read(file.path(preprocessed_greenland_data_path,"shrunk300_scoresby_sound_utm.shp"))
shrunk300_scoresby_sound_utm <- st_transform(shrunk300_scoresby_sound_utm, crs = "+init=EPSG:32626 +units=km")

@@

```{r compute updated translated distance to shore}
res_shrunk100_land=nearest_points(positions=narwhal_positions,border=shrunk100_scoresby_sound_utm)
res_shrunk200_land=nearest_points(positions=narwhal_positions,border=shrunk200_scoresby_sound_utm)
res_shrunk300_land=nearest_points(positions=narwhal_positions,border=shrunk300_scoresby_sound_utm)

cat("Percentage of points on land with distance from 100m shrunk land polygons : \n",
    length(which(res_shrunk100_land$min_distances==0))/length(dataprep$time)*100,"\n")

cat("Percentage of points on land with distance from 200m shrunk land polygons : \n",
    length(which(res_shrunk200_land$min_distances==0))/length(dataprep$time)*100,"\n")

cat("Percentage of points on land with distance from 300m shrunk land polygons : \n",
    length(which(res_shrunk300_land$min_distances==0))/length(dataprep$time)*100,"\n")

@@

```{r distribution of points close to the shore for shrunk polygons}

df100=data.frame(DistanceShore=res_shrunk100_land$min_distances[res_shrunk100_land$min_distances<1.5])
df200=data.frame(DistanceShore=res_shrunk200_land$min_distances[res_shrunk200_land$min_distances<1.5])
df300=data.frame(DistanceShore=res_shrunk300_land$min_distances[res_shrunk300_land$min_distances<1.5])

df100$Category="100m"
df200$Category="200m"
df300$Category="300m"

df<- rbind(df100,df200,df300)
histo=ggplot(df, aes(x =DistanceShore,fill=Category)) +
  geom_histogram(bins=100,color="black")+xlab("Computed distance to shore (km)")

@@

add_boundary_data<-function(
    data,res,names=list(px="px",py="py",DistanceShore="DistanceShore",nx="nx",ny="ny"))

  {

  #data : dataframe to which wa add columns 
  # res : list as returned by nearest points, with elements
  # nearest points and min_distances
  # names for the distance, nearest points coordinates and normal vectors

  nearest_points=res$nearest_points
  min_distances=res$min_distances

  #add columns px and py with coordinates of the nearest point on the shore line
  data[[names$px]]=nearest_points[,1]
  data[[names$py]]=nearest_points[,2]

  # add columns with difference between narwhal position and nearest point on the shoreline
  data[[names$nx]]=data$x- data[[names$px]]
  data[[names$ny]]=data$y-data[[names$py]]

  #replace distance to shore with distances calculated from land polygons (in km)
  data[[names$DistanceShore]]=min_distances

  return(data)
}

dataprep<- add_boundary_data(dataprep,res_shrunk100_land,
                  names=list(px="px100",py="py100",DistanceShore="DistanceShore100",nx="nx100",ny="ny100"))

dataprep <- add_boundary_data(dataprep,res_shrunk200_land,
                  names=list(px="px200",py="py200",DistanceShore="DistanceShore200",nx="nx200",ny="ny200"))

dataprep <- add_boundary_data(dataprep,res_shrunk300_land,
                  names=list(px="px300",py="py300",DistanceShore="DistanceShore300",nx="nx300",ny="ny300"))

@@

```{r}
#add Ishore covariate based on 1/distance in km
dataprep$Ishore100 <- ExposureFunction(X = dataprep$DistanceShore100) 
dataprep$Ishore200 <- ExposureFunction(X = dataprep$DistanceShore200) 
dataprep$Ishore300 <- ExposureFunction(X = dataprep$DistanceShore300)

@@

dataprep$AngleVelocities=rep(NA,n)
dataprep$theta100=rep(NA,n)
dataprep$theta200=rep(NA,n)
dataprep$theta300=rep(NA,n)
dataprep$VelocityPhase=rep(NA,n)

@@

  #matrix of normal vectors
  normal100=as.matrix(sub_data[2:n_sub,c("nx100","ny100")])
  normal200=as.matrix(sub_data[2:n_sub,c("nx200","ny200")])
  normal300=as.matrix(sub_data[2:n_sub,c("nx300","ny300")])

  #angle between velocity and normal vector
  theta100=signed_angle(vexp_df,normal100)
  theta200=signed_angle(vexp_df,normal200)
  theta300=signed_angle(vexp_df,normal300)

@@

  #adjust lengths by adding NA when necessary
  angle_vel=c(NA,angle_vel,NA)
  theta100=c(NA,theta100)
  theta200=c(NA,theta200)
  theta300=c(NA,theta300)
  vel_phase=c(NA,vel_phase,NA)

  dataprep[sub_ind,"AngleVelocities"]=angle_vel
  dataprep[sub_ind,"theta100"]=-theta100
  dataprep[sub_ind,"theta200"]=-theta200
  dataprep[sub_ind,"theta300"]=-theta300
  dataprep[sub_ind,"VelocityPhase"]=vel_phase
}

@@

@@

plot_AngleNormal=ggplot()+geom_line(data=dataprep[dataprep$DistanceShore>0.05,],aes(time,theta100,col=ID))+
  geom_rect(data = regions, aes(xmin = start_time, xmax = end_time, ymin = -Inf, ymax = Inf),alpha = 0.3,col="lightblue")+
  facet_grid(ID ~.,labeller=as_labeller(c("Asgeir"="A1","Frederik"="A2","Helge18"="A3","Kyrri"="A4","Nemo"="A5","Siggi"="A6")))+
  scale_color_viridis_d(guide="none")+theme(legend.position="none")+xlab("Time (h)")+ylab("Angle between empirical velocity and normal vector")+ theme_minimal()

ggsave(path=plots_folder,"plot_horizontal_speed.png", plot = plot_speed)
ggsave(path=plots_folder,"plot_AngleVelocities.png", plot = plot_AngleVelocities)
ggsave(path=plots_folder,"plot_AngleNormal.png", plot = plot_AngleNormal)

@@

@@

@@

AngleNormal_histo=ggplot(dataprep[dataprep$DistanceShore>0.05,], aes(x = theta100,col=ID)) +
  geom_histogram(fill="white",bins=50)+facet_grid(ID ~.,labeller=as_labeller(c("Asgeir"="A1","Frederik"="A2","Helge18"="A3","Kyrri"="A4","Nemo"="A5","Siggi"="A6")))+
  scale_color_viridis_d(guide="none")+
  theme(legend.position="top")+xlab("Angle between empirical velocity and normal vector (rad) ")+ theme_minimal()

@@

ggsave(path=plots_folder,"AngleVelocities.png", plot = AngleVelocities_histo)
ggsave(path=plots_folder,"AngleNormal.png", plot = AngleNormal_histo)
ggsave(path=plots_folder,"VelocityPhase.png", plot = VelocityPhase_histo)
```

@@

@@

@@

@@

```{r plot empirical velocities and deviation angles without points on land}

@@

```{r distance shore histogram}
Dshore100_histo=ggplot(dataprep, aes(x = DistanceShore100)) +
  geom_histogram(fill="white",color="black",bins=50)+
  facet_grid(ID ~.,labeller=as_labeller(c("Asgeir"="A1","Frederik"="A2","Helge18"="A3","Kyrri"="A4","Nemo"="A5","Siggi"="A6")))+
  theme(legend.position="none")+xlab("Distance to shore (km)")

Dshore200_histo=ggplot(dataprep, aes(x = DistanceShore200)) +
  geom_histogram(fill="white",color="black",bins=50)+
  facet_grid(ID ~.,labeller=as_labeller(c("Asgeir"="A1","Frederik"="A2","Helge18"="A3","Kyrri"="A4","Nemo"="A5","Siggi"="A6")))+
  theme(legend.position="none")+xlab("Distance to shore (km)")

Dshore300_histo=ggplot(dataprep, aes(x = DistanceShore300)) +
  geom_histogram(fill="white",color="black",bins=50)+
  facet_grid(ID ~.,labeller=as_labeller(c("Asgeir"="A1","Frederik"="A2","Helge18"="A3","Kyrri"="A4","Nemo"="A5","Siggi"="A6")))+
  theme(legend.position="none")+xlab("Distance to shore (km)")

ggsave(path=plots_folder,"Dshore100_histo.png", plot = Dshore100_histo)
ggsave(path=plots_folder,"Dshore200_histo.png", plot = Dshore200_histo)
ggsave(path=plots_folder,"Dshore300_histo.png", plot = Dshore300_histo)
```

```{r display Dshore_histo, fig.cap="Histograms of distance to shore"}
Dshore100_histo
Dshore200_histo
Dshore300_histo
```
## Plot distance to ship histogram for each narwhal

@@

Dship_histo=ggplot(dataprep, aes(x = Dist_Ship)) +
  geom_histogram(fill="white",col="black",bins=50)+
  facet_grid(ID ~.,labeller=as_labeller(c("Asgeir"="A1","Frederik"="A2","Helge18"="A3","Kyrri"="A4","Nemo"="A5","Siggi"="A6")))+
  theme(legend.position="none")+xlab("Distance to ship (km)")

@@

```{r}
D_low=0.07
D_up=3
df=cropped_data
df[df$DistanceShore<D_low,"Ishore100"]=1/D_low
df[df$DistanceShore>D_up,"Ishore100"]=0

pIshore=ggplot(df,aes(time,Ishore100,col=ID))+ geom_line(color="black")+facet_grid(ID ~.,labeller=as_labeller(c("Asgeir"="A1","Frederik"="A2","Helge18"="A3","Kyrri"="A4","Nemo"="A5","Siggi"="A6")))+
  theme(legend.position="none")+xlab("Time (h)")+ylab("Inverse distance to shore")

ggsave(path=plots_folder,"realIshore100_through_time.png", plot=pIshore)
```

@@

# Animate trajectories with Ship Exposure

```{r simulate response trajectories with land reflection,cache=TRUE,message=FALSE}

# Plot both in a single plot with facets
ggplot(dataprep) +geom_sf(data=shrunk100_scoresby_sound_utm$geometry,fill="grey")+
  coord_sf(datum=st_crs("+init=EPSG:32626 +units=km"))+
  geom_point(aes(x = x, y = y, color = ID), size = 0.6) + 
  ggtitle("Trajectories during ship exposure") +
  theme_minimal() +
  labs(color = "Narwhal ID")

```{r animate response trajectories with land reflection, results=FALSE,eval=interactive()}

# Basic ggplot setup
plot_trajectories_land <- ggplot() +
  geom_sf(data=shrunk100_scoresby_sound_utm$geometry,fill="grey")+
  coord_sf(datum=st_crs("+init=EPSG:32626 +units=km"))+
  geom_path(data=dataprep, 
            aes(x = x, y = y, color =ExpShip,group=ID),size = 1, 
            lineend = "round",alpha=0.15) +     
  geom_point(data=dataprep,
             aes(x = x, y = y, color =ExpShip,group=ID,shape=ID),size =4 ) +  
  scale_color_gradient(low = "blue", high = "red") +
  labs(title = "Narwhal trajectories with ship exposure", 
       subtitle = "Time : {frame_along} h", 
       x = "X Position", 
       y = "Y Position") +
  theme_minimal()

anim <- plot_trajectories_land + 
  transition_reveal(along = Posixct) + 
  labs(subtitle = "Time: {round(frame_along,1)} hours")

# Save the animation
animate(anim, nframes = 100, fps = 5, width = 800, height = 600, 
        renderer = gifski_renderer("trajectories_ExpShip.gif"))
